---
title: "A look at the US top 10 most visited National Parks in 2020"
description: |
  Talia Chait, Erin Franke, Alex Ang, Katie Funk
site: distill::distill_website
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, include= FALSE}
library(readr)
library(tidyverse)
library(stringr)
library(tidyr)
library(ggridges)
library(nationalparkcolors)
library(ggthemes) 
library(ggtext)
library(patchwork)
library(gt)
nationalpark <- read_csv("data/nationalpark.csv")
`%!in%` <- Negate(`%in%`)

national_parks <- nationalpark %>%
  mutate(lat = as.numeric(substr(`_geoloc`, 9, 16)), 
         long = as.numeric(substr(`_geoloc`, 26,33)), 
         length_miles = round(length/5280, 3), 
         ) %>%
  select(name, area_name, lat, long, city_name, state_name, popularity, length_miles, elevation_gain, difficulty_rating, route_type, visitor_usage, avg_rating, num_reviews, features, activities) %>%
  filter(area_name %in% c("Great Smoky Mountains National Park", "Yellowstone National Park", "Zion National Park", "Rocky Mountain National Park", "Grand Teton National Park", "Grand Canyon National Park", "Cuyahoga Valley National Park", "Acadia National Park", "Olympic National Park", "Joshua Tree National Park")) %>%
  mutate(no_dogs = as.numeric(str_detect(features, "dogs-no")), 
         river = as.numeric(str_detect(features, "river")), 
         views = as.numeric(str_detect(features, "views")), 
         wildflowers = as.numeric(str_detect(features, "wild-flowers")), 
         wildlife = as.numeric(str_detect(features, "wildlife")), 
         cave = as.numeric(str_detect(features, "cave")), 
         forest = as.numeric(str_detect(features, "forest")), 
         partially_paved = as.numeric(str_detect(features, "partially-paved")), 
         ada = as.numeric(str_detect(features, "ada")), 
         waterfall = as.numeric(str_detect(features, "waterfall")), 
         kids = as.numeric(str_detect(features, "kids")), 
         paved = as.numeric(str_detect(features, "paved")), 
         beach = as.numeric(str_detect(features, "beach")), 
         strollers = as.numeric(str_detect(features, "strollers")), 
         dogs_leash = as.numeric(str_detect(features, "dogs-leash")), 
         historic_site = as.numeric(str_detect(features, "historic-site")), 
         lake = as.numeric(str_detect(features, "lake")), 
         hiking = as.numeric(str_detect(activities, "hiking")), 
         nature_trips = as.numeric(str_detect(activities, "nature-trips")), 
         trail_running = as.numeric(str_detect(activities, "trail-running")), 
         birding = as.numeric(str_detect(activities, "birding")), 
         fishing = as.numeric(str_detect(activities, "fishing")), 
         walking = as.numeric(str_detect(activities, "walking")), 
         backpacking = as.numeric(str_detect(activities, "backpacking")), 
         camping = as.numeric(str_detect(activities, "camping")), 
         road_biking = as.numeric(str_detect(activities, "road-biking")), 
         rock_climbing = as.numeric(str_detect(activities, "rock-climbing")), 
         cross_country_skiing = as.numeric(str_detect(activities, "cross-country-skiing")), 
         paddle_sports = as.numeric(str_detect(activities, "paddle-sports")), 
         snowshoeing = as.numeric(str_detect(activities, "snowshoeing")), 
         fly_fishing = as.numeric(str_detect(activities, "fly-fishing")), 
         horseback_riding = as.numeric(str_detect(activities, "horseback-riding")), 
         mountain_biking = as.numeric(str_detect(activities, "mountain-biking")), 
         scenic_driving = as.numeric(str_detect(activities, "scenic-driving")), 
         bike_touring = as.numeric(str_detect(activities, "bike-touring")), 
         off_road_driving = as.numeric(str_detect(activities, "off-road-driving")))
national_parks$total_features <- rowSums(national_parks[17:52])

kids_proportion <- national_parks %>%
  filter(name %!in% c("Olympic Peninsula Loop Drive", "West Yellowstone to Jackson Scenic Drive", "Joshua Tree National Park Scenic Drive")) %>%
  group_by(area_name, kids) %>%
  count() %>%
  pivot_wider(id_cols = area_name, names_from = kids, values_from = n) %>%
  rename(kids_no = `0`, kids_yes = `1`) %>%
  mutate(prop_kid_friendly = round(kids_yes / (kids_yes + kids_no),3)) %>%
  arrange(desc(prop_kid_friendly))

kids_rating <- national_parks %>%
  filter(name %!in% c("Olympic Peninsula Loop Drive", "West Yellowstone to Jackson Scenic Drive", "Joshua Tree National Park Scenic Drive"), 
         kids == 1) %>%
  group_by(area_name) %>%
  summarize(kids_avg_rating= mean(avg_rating, na.rm = TRUE))

kids_proportion <- kids_proportion %>%
  inner_join(kids_rating, by = "area_name")
```

### The top 10 most visited national parks in 2020

Here are the parks that we looked at in our analysis. 

```{r, echo = FALSE}
national_parks %>%
  filter(area_name %in% c("Great Smoky Mountains National Park", "Yellowstone National Park", "Zion National Park", "Rocky Mountain National Park", "Grand Teton National Park", "Grand Canyon National Park", "Cuyahoga Valley National Park", "Acadia National Park", "Olympic National Park", "Joshua Tree National Park")) %>%
  distinct(area_name) %>%
  mutate(index  = c(6, 10, 4, 8, 2, 1, 7, 3, 9, 5)) %>%
  arrange(index) %>%
  select(area_name) %>%
  rename(Park = area_name) %>%
  gt() %>%
  cols_label(
    Park = md("**Top 10 most visited US National Parks**")) %>%
  tab_style(
    style = cell_fill(color = "lightcyan2"),
    locations = cells_body(
      rows = Park %in% c("Great Smoky Mountains National Park", "Zion National Park", "Grand Teton National Park", "Olympic National Park", "Cuyahoga Valley National Park"))) %>%
  tab_style(
    style = cell_fill(color = "lightgoldenrodyellow"),
    locations = cells_body(
      rows = Park %!in% c("Great Smoky Mountains National Park", "Zion National Park", "Grand Teton National Park", "Olympic National Park", "Cuyahoga Valley National Park")))
```

### Trips to the National Parks with kids

A final national parks topic we decided to focus on was the kid friendliness aspect of the national parks. Some trails, such as the famous Angel’s Landing at Zion National Park, have steep cliffs and are incredibly dangerous for young children. To help families more easily choose appropriate hiking trails on their trips, we summarized the data based on what trails has a feature listed as “kids”. The graph below shows the distribution of hiking distances for ten mile or less trails that are recommended as kid friendly versus not kid friendly. On the other hand, at all ten parks the large majority of kid friendly hiking trails are less than three miles. Out of all the parks, it seems that Acadia tends to have the highest proportion of less than three mile hikes (both kid and not kid friendly), with very few hikes in the 4-10 mile range.

```{r, echo = FALSE}
national_parks %>%
  filter(name %!in% c("Olympic Peninsula Loop Drive", "West Yellowstone to Jackson Scenic Drive", "Joshua Tree National Park Scenic Drive")) %>%
  left_join(kids_proportion, by = "area_name") %>%
  mutate(area_name = as.factor(area_name), 
         area_name = word(area_name, 1, -3)) %>%
  ggplot(aes(y=fct_reorder(area_name, kids_yes)))+
  geom_density_ridges(
    aes(x=length_miles, fill = paste(area_name, kids)), 
    alpha = .8, color = "white", from = 0, to =40)+
  scale_y_discrete(expand = c(0, 0)) +
  scale_x_continuous(breaks = c(0, 2, 4,6, 8, 10), limits = c(0, 10), expand = c(0, 0)) +
  scale_fill_cyclical(
    breaks = c(0, 1),
    labels = c(`0` = "no_kids", `1` = "yes_kids"),
    values = c("salmon2", "turquoise3", "salmon2", "turquoise3"),
    name = "Option", guide = "legend"
  ) +
  labs(x= "", y="", title = "Trail length (miles) for <strong><span style='color:turquoise3'>kid</span></strong></b> versus <strong><span style='color:salmon2'>not kid</span></strong></b> recommended trails")+
  theme_minimal()+
  theme(plot.title.position = "plot", 
        panel.grid.major.x = element_blank(), 
        panel.grid.minor.x = element_blank(),
        plot.title = element_markdown(family = "mono", face = "bold", size = 10))
```

After getting the sense that kid friendly trails tend to be shorter, we decided to look more into each park and see if there might be one that is more attractive to familie than another. To do this, we looked at the number of kid friendly trails and the average rating those trails got. The results are summarized below. 

```{r, echo = FALSE}
kids_proportion %>%
  mutate(area_name = word(area_name, 1, -3)) %>%
  ggplot(aes(y=fct_reorder(area_name, kids_yes), x=kids_yes))+
  geom_col(aes(fill = kids_avg_rating)) + 
  labs(y="", x="", 
       title = "Number of trails that are <strong><span style='color:dodgerblue3'>kid friendly</span></strong></b>",
       fill = "Avg <strong><span style='color:dodgerblue3'>kid friendly</span></strong></b> \ntrail rating")+
  scale_x_continuous(expand = c(0,0))+
  theme_minimal()+
  theme(plot.title = element_markdown(family = "mono", face = "bold"),
        legend.title = element_markdown(family = "mono", size = 7),
        plot.title.position = "plot", 
        panel.grid.major.x = element_blank(), 
        panel.grid.minor.x = element_blank(), 
        panel.grid.major.y = element_blank(), 
        panel.grid.minor.y = element_blank())
```

From this graph we can see both Yellowstone and Acadia have close to 80 kid recommended trails...it might take quite a few family vacations to cover all of those :) On the lower end are Grand Teton and Grand Canyon National with 24 and 27 kid recommended trails each. The graph is colored by average kid friendly trail rating. While Grand Teton has the fewest number of kid friendly trails - though still penty for a family vacation with over 52 miles of these trails - the average trail rating is the highest. Joshua Tree has the lowest average kid recommended trail rating at around 3.8/5. 

To wrap up our discussion of the national parks and kid friendly trails, we recommended two family friendly trails from each of the ten parks we looked at. To choose these parks, we looked for trails within our data that have a high number of reviews and at least 4.5 stars. The table below highlights some details of these trails including a few of their unique features. 

```{r, echo = FALSE}
recommend_trails <- national_parks %>%
  filter(name %in% c("Peregrine Peak via Alum Cave Bluffs Trail", 
                     "Clingmans Dome Observation Tower Trail", 
                     "Grand Prismatic Hot Spring", 
                     "Mystic Falls, Fairy Creek and Little Firehole Loop", 
                     "Zion Canyon Overlook Trail", 
                     "The Watchman Trail", 
                     "Emerald Lake Trail", 
                     "Dream Lake Trail", 
                     "Hidden Falls Trail", 
                     "String Lake Trail", 
                     "Shoshone Point Trail", 
                     "Grand Canyon Rim Trail", 
                     "Ledges Trail", 
                     "Brandywine Gorge Trail", 
                     "Cadillac North Ridge Trail", 
                     "Jordan Pond Full Loop Trail", 
                     "Hoh Rain Forest Hall of Moss", 
                     "Marymere Falls Trail", 
                      "Ryan Mountain Trail", 
                    "Hidden Valley Nature Trail")) %>%
  select(area_name, name, length_miles, avg_rating, num_reviews)%>%
  head(20) %>%
  mutate(`noteable features` = c("views, wildlife", "views, wildlife", "hot springs, river", "hot springs, river, waterfall", "cave, trail running", "river, camping", "river, fishing, snowshoeing", "skiing, snowshoeing, rock climbing", "lake, beach, rock climbing, camping", "ada, beach, strollers, birding", "ada, strollers, birding, trail running", "ada, strollers", "historic site, trail running", "historic site, waterfall, river, fishing", "trail running, birding, wildlife", "lake, birding, trail running", "ada, strollers, camping, birding", "lake, waterfall, birding", "wildlife, views", "views, rock climbing, trail running")) %>%
  rename(Park = area_name, Trail = name, `Length (miles)` = length_miles, `Avg rating` = avg_rating, `Number reivews` = num_reviews, `Noteable features` = `noteable features`) %>%
  arrange(Park, Trail)

trail_recs_table <- recommend_trails %>%
  gt() %>%
  tab_header(
    title = md("**Exciting trails for kids**"),
    subtitle = md("**Two kid recommended trails from each of the top 10 most visited national parks**")
  ) %>%
  tab_style(
    style = cell_fill(color = "lightgoldenrodyellow"),
    locations = cells_body(
      rows = Park %in% c("Acadia National Park", "Grand Canyon National Park", "Great Smoky Mountains National Park", "Olympic National Park", "Yellowstone National Park"))) %>%
  tab_style(
    style = cell_fill(color = "lightcyan2"),
    locations = cells_body(
      rows = Park %!in% c("Acadia National Park", "Grand Canyon National Park", "Great Smoky Mountains National Park", "Olympic National Park", "Yellowstone National Park"))) %>%
  tab_options(
    table.font.size = px(13L)
  )
trail_recs_table
```

We hope you enjoy your next trip to one of the many beautiful US national parks! 
