---
title: "A look at the US top 10 most visited National Parks in 2020"
description: |
  Talia Chait, Erin Franke, Alex Ang, Katie Funk
site: distill::distill_website
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, include= FALSE}
library(readr)
library(tidyverse)
library(stringr)
library(tidyr)
library(ggridges)
library(nationalparkcolors)
library(ggthemes) 
library(ggtext)
library(patchwork)
library(gt)
library(maps)
library(openintro)
library(leaflet)
library(lubridate)     # for date manipulation
library(openintro)     # for the abbr2state() function
library(palmerpenguins)# for Palmer penguin data
library(maps)          # for map data
library(ggmap)         # for mapping points on maps
library(gplots)        # for col2hex() function
library(RColorBrewer)  # for color palettes
library(sf)            # for working with spatial data
library(leaflet)       # for highly customizable mapping
library(carData)       # for Minneapolis police stops data
library(ggtext)
theme_set(theme_minimal())
nationalpark <- read_csv("data/nationalpark.csv")
`%!in%` <- Negate(`%in%`)

national_parks <- nationalpark %>%
  mutate(lat = as.numeric(substr(`_geoloc`, 9, 16)), 
         long = as.numeric(substr(`_geoloc`, 26,33)), 
         length_miles = round(length/5280, 3), 
         ) %>%
  select(name, area_name, lat, long, city_name, state_name, popularity, length_miles, elevation_gain, difficulty_rating, route_type, visitor_usage, avg_rating, num_reviews, features, activities) %>%
  filter(area_name %in% c("Great Smoky Mountains National Park", "Yellowstone National Park", "Zion National Park", "Rocky Mountain National Park", "Grand Teton National Park", "Grand Canyon National Park", "Cuyahoga Valley National Park", "Acadia National Park", "Olympic National Park", "Joshua Tree National Park")) %>%
  mutate(no_dogs = as.numeric(str_detect(features, "dogs-no")), 
         river = as.numeric(str_detect(features, "river")), 
         views = as.numeric(str_detect(features, "views")), 
         wildflowers = as.numeric(str_detect(features, "wild-flowers")), 
         wildlife = as.numeric(str_detect(features, "wildlife")), 
         cave = as.numeric(str_detect(features, "cave")), 
         forest = as.numeric(str_detect(features, "forest")), 
         partially_paved = as.numeric(str_detect(features, "partially-paved")), 
         ada = as.numeric(str_detect(features, "ada")), 
         waterfall = as.numeric(str_detect(features, "waterfall")), 
         kids = as.numeric(str_detect(features, "kids")), 
         paved = as.numeric(str_detect(features, "paved")), 
         beach = as.numeric(str_detect(features, "beach")), 
         strollers = as.numeric(str_detect(features, "strollers")), 
         dogs_leash = as.numeric(str_detect(features, "dogs-leash")), 
         historic_site = as.numeric(str_detect(features, "historic-site")), 
         lake = as.numeric(str_detect(features, "lake")), 
         hiking = as.numeric(str_detect(activities, "hiking")), 
         nature_trips = as.numeric(str_detect(activities, "nature-trips")), 
         trail_running = as.numeric(str_detect(activities, "trail-running")), 
         birding = as.numeric(str_detect(activities, "birding")), 
         fishing = as.numeric(str_detect(activities, "fishing")), 
         walking = as.numeric(str_detect(activities, "walking")), 
         backpacking = as.numeric(str_detect(activities, "backpacking")), 
         camping = as.numeric(str_detect(activities, "camping")), 
         road_biking = as.numeric(str_detect(activities, "road-biking")), 
         rock_climbing = as.numeric(str_detect(activities, "rock-climbing")), 
         cross_country_skiing = as.numeric(str_detect(activities, "cross-country-skiing")), 
         paddle_sports = as.numeric(str_detect(activities, "paddle-sports")), 
         snowshoeing = as.numeric(str_detect(activities, "snowshoeing")), 
         fly_fishing = as.numeric(str_detect(activities, "fly-fishing")), 
         horseback_riding = as.numeric(str_detect(activities, "horseback-riding")), 
         mountain_biking = as.numeric(str_detect(activities, "mountain-biking")), 
         scenic_driving = as.numeric(str_detect(activities, "scenic-driving")), 
         bike_touring = as.numeric(str_detect(activities, "bike-touring")), 
         off_road_driving = as.numeric(str_detect(activities, "off-road-driving")))
national_parks$total_features <- rowSums(national_parks[17:52])

kids_proportion <- national_parks %>%
  filter(name %!in% c("Olympic Peninsula Loop Drive", "West Yellowstone to Jackson Scenic Drive", "Joshua Tree National Park Scenic Drive")) %>%
  group_by(area_name, kids) %>%
  count() %>%
  pivot_wider(id_cols = area_name, names_from = kids, values_from = n) %>%
  rename(kids_no = `0`, kids_yes = `1`) %>%
  mutate(prop_kid_friendly = round(kids_yes / (kids_yes + kids_no),3)) %>%
  arrange(desc(prop_kid_friendly))

kids_rating <- national_parks %>%
  filter(name %!in% c("Olympic Peninsula Loop Drive", "West Yellowstone to Jackson Scenic Drive", "Joshua Tree National Park Scenic Drive"), 
         kids == 1) %>%
  group_by(area_name) %>%
  summarize(kids_avg_rating= mean(avg_rating, na.rm = TRUE))

kids_proportion <- kids_proportion %>%
  inner_join(kids_rating, by = "area_name")
```

### What were the top 10 most visited national parks in 2020?

Here are the parks that we looked at in our analysis. 

```{r, echo = FALSE}
national_parks %>%
  filter(area_name %in% c("Great Smoky Mountains National Park", "Yellowstone National Park", "Zion National Park", "Rocky Mountain National Park", "Grand Teton National Park", "Grand Canyon National Park", "Cuyahoga Valley National Park", "Acadia National Park", "Olympic National Park", "Joshua Tree National Park")) %>%
  distinct(area_name) %>%
  mutate(index  = c(6, 10, 4, 8, 2, 1, 7, 3, 9, 5)) %>%
  arrange(index) %>%
  select(area_name) %>%
  rename(Park = area_name) %>%
  gt() %>%
  cols_label(
    Park = md("**Top 10 most visited US National Parks**")) %>%
  tab_style(
    style = cell_fill(color = "lightcyan2"),
    locations = cells_body(
      rows = Park %in% c("Great Smoky Mountains National Park", "Zion National Park", "Grand Teton National Park", "Olympic National Park", "Cuyahoga Valley National Park"))) %>%
  tab_style(
    style = cell_fill(color = "lightgoldenrodyellow"),
    locations = cells_body(
      rows = Park %!in% c("Great Smoky Mountains National Park", "Zion National Park", "Grand Teton National Park", "Olympic National Park", "Cuyahoga Valley National Park")))
```


Of the top 10 national parks, first we first made a quick overview of all the trails together to give you a sense of the differences in length and elevation. As one can see, for the length (which is in miles) the majority of all the trails are between 0-10 miles long with some outsiders of more than 10 miles. As for elevation, the majority of all the trails have an elevation gain between 0 and 3,000 feet.


```{r, echo = FALSE}
plot1 <- national_parks %>% 
  group_by(area_name) %>% 
  ggplot(aes(x = length_miles)) + 
  geom_density(fill = "salmon2") +
  scale_x_continuous(breaks = c(0, 10, 20, 30, 40, 50, 60), 
                     limits = c(0, 65), 
                     expand = c(0, 0)) +
  labs(x = "",
       y = "",
       title = "**Density Plot of Different Trail Lengths (in miles)**",
       subtitle = "<span style='font-size:11pt'> For Top 10 National Parks",
       caption = "") +
  theme(plot.title.position = "plot",
        plot.title = element_markdown(family = "mono"),
        plot.subtitle = element_markdown(family = "mono")) +
  geom_segment(aes(x = 12, y = 0.20, xend = 2, yend = 0.23),
             arrow = arrow(length = unit(0.03, "npc"))) +
  annotate(geom = "text", 
           y = 0.15,
           x = 12, 
           label = "The majority of trail lengths 
                     are between 0 and 10 miles!", 
           cex = 3)

plot2 <- national_parks %>% 
  group_by(area_name) %>% 
  ggplot(aes(x = elevation_gain)) + 
  geom_density(fill = "turquoise3") +
  scale_x_continuous(breaks = c(0, 1000, 2000, 3000, 4000, 5000, 6000), 
                     limits = c(0, 6500), 
                     expand = c(0, 0)) +
  labs(x = "",
       y = "",
       title = "**Density Plot of Different Trail Elevations (in feet)**",
       subtitle = "<span style='font-size:11pt'> For Top 10 National Parks",
       caption = "") +
  theme(plot.title.position = "plot",
        plot.title = element_markdown(family = "mono"),
        plot.subtitle = element_markdown(family = "mono")) 

plot1 + plot2 +
   plot_layout(ncol = 1, heights = c(5, 5))
```

### Where are these national parks located?

Visually, we thought it would be helpful to plot the trails by park on a map of the United States. We can see the distribution of parks and their relative location on the US Map. After someone decides which park they want to visit, they can zoom in to find an interactive map showing the varying lengths (miles) of each trail within the top 10 parks. You can also hover over each of the trails to see them labeled by trail name. 

```{r, echo = FALSE}
unitedstates <- get_stamenmap(
    bbox = c(left = -130.15, bottom = 19.10, right = -60.64, top = 55.77), 
    maptype = "terrain",
    zoom = 4)

#ggmap(unitedstates) +
#  geom_point(data=national_parks, aes(x=long, y=lat, size=length_miles, color=area_name), alpha=.3) +
 # theme(text = element_text(family = "mono", face = "bold"))

pal <- colorFactor(palette="viridis", domain=national_parks$area_name)


leaflet(data=national_parks) %>% 
  setMaxBounds( lng1 = -145.22
                , lat1 = 58.08
                , lng2 = -51.01
                , lat2 = 8.66 ) %>%
  addProviderTiles(providers$Stamen.TonerLite) %>% 
  addCircles(lng= ~long,
           lat = ~lat, 
           color= ~pal(area_name),
           label= ~name,
           radius= ~length_miles) %>%
  addLegend(position = "bottomright",
            pal = pal, 
            values = ~area_name,
            title = "Top Ten Visited Parks in 2020", 
            )
```

### Which parks are the most popular?

In the United States, there are numerous parks someone can visit, but given time constraints, they might want to consider traveling to the most popular National Parks in the US. These parks are located all across the United States, and each has different amenities, and views to offer. Rocky Mountain National Park in Colorado(the most popular park in the US) offers mountainous views and hiking while Joshua Tree National Park in California gives its viewers the opportunity to wander through beautiful desert landscapes. 

To ensure visitors get the best experience while visiting their National Park of choice, and get the opportunity to see the best sites, we found the hiking trails at each national park that have the most visitors. While these might not be everyone’s cup of tea(more touristy attractions than off the beaten path trails), they can provide casual campers with unforgettable experiences.

```{r, echo = FALSE}
top_park_location <- national_parks %>%
  group_by(area_name) %>%
  filter(popularity == max(popularity))
  

national_parks %>%
  group_by(area_name) %>%
  summarise(total_park_pop = sum(popularity)) %>%
  ggplot(aes(y = fct_reorder(area_name, total_park_pop), x = total_park_pop)) +
  geom_col(fill = "salmon2") +
  geom_col(data = top_park_location, aes(y = area_name, x = popularity),
           fill = "gold2") +
  geom_text(data = top_park_location, aes(y = area_name, x = popularity,
                                          label = name),hjust = 0) +
  theme(panel.grid.major.y = element_blank(),
        plot.title = element_markdown(family = "mono", face = "bold"),
        plot.title.position = "plot") +
  labs(title = "Visitation at the Ten Most Popular National Parks",
       y = "",
       x = "")
```

### What are the different route types available? 

On top of the prior aspects that we analyzed, we also decided to look into the route type of all the trails at the different national parks. Some trails are easy to follow when hiking while others require one to know where they are going. There are three route type categories: out and back, loop, and point to point. The graph below shows the number of trails per route type at each national park.

```{r, echo = FALSE}
national_parks %>%
  group_by(area_name, route_type) %>% 
  summarize(nroutetype = n()) %>% 
  ungroup() %>% 
  ggplot(mapping = aes(x = fct_reorder(area_name, nroutetype, sum, .desc = FALSE), 
                       y = nroutetype,
                       fill = route_type)) + 
  geom_col(position = "stack") +
  coord_flip() + 
  scale_fill_manual(values = c("turquoise3", "salmon2", "gold2")) +
  labs(x = "",
       y = "",
       title = "**Total Number of Trails in the Top 10 National Parks by Route Type**",
       subtitle = "<span style='font-size:10pt'> Route types are **<span style='color:gold2;'>point to point</span>**, **<span style='color: salmon2;'>out and back</span>**, and **<span style = 'color: turquoise3;'>loop</span>**",
       fill = "",
       caption = "") +
  theme(legend.position = "none",
        plot.title.position = "plot",
        plot.title = element_markdown(family = "mono", size = 11),
        plot.subtitle = element_markdown(family = "mono"))
```

As one can see, the most popular route types are out and back and loop for all 10 parks. Specifically the out and back is likely more popular than the loop. Most of the parks have at least 50 + out and back trails with some of the larger parks having closer to 100+ out and back trails. While some route types may seem easier to hike than others, it is also important to look at the difficulty of the trail, especially if you have younger kids or are a less experienced hiker! 

### Are there trails or parks that are more kid friendly than others?

When you are planning a hiking trip with kids, you will probably want to do everything in your power to make it as painful as possible. In order to do this, we recommend going on trails that are more family friendly. Some trails, such as the famous Angel’s Landing at Zion National Park, have steep cliffs and are incredibly dangerous for young children. To help families more easily choose appropriate hiking trails on their trips, we summarized the data based on what trails has a feature listed as “kids”. In the graph below, we can see the distribution of hiking distances for ten mile or less trails that are classified as kid friendly versus not kid friendly. At all ten parks the large majority of kid friendly hiking trails are less than three miles, which makes sense. It seems that out of all the parks that Acadia tends to have the highest proportion of less than three mile hikes (both kid and not kid friendly), with very few hikes in the 4-10 mile range.

```{r, echo = FALSE}
national_parks %>%
  filter(name %!in% c("Olympic Peninsula Loop Drive", "West Yellowstone to Jackson Scenic Drive", "Joshua Tree National Park Scenic Drive")) %>%
  left_join(kids_proportion, by = "area_name") %>%
  mutate(area_name = as.factor(area_name), 
         area_name = word(area_name, 1, -3)) %>%
  ggplot(aes(y=fct_reorder(area_name, kids_yes)))+
  geom_density_ridges(
    aes(x=length_miles, fill = paste(area_name, kids)), 
    alpha = .8, color = "white", from = 0, to =40)+
  scale_y_discrete(expand = c(0, 0)) +
  scale_x_continuous(breaks = c(0, 2, 4,6, 8, 10), limits = c(0, 10), expand = c(0, 0)) +
  scale_fill_cyclical(
    breaks = c(0, 1),
    labels = c(`0` = "no_kids", `1` = "yes_kids"),
    values = c("salmon2", "turquoise3", "salmon2", "turquoise3"),
    name = "Option", guide = "legend"
  ) +
  labs(x= "", y="", title = "Trail length (miles) for <strong><span style='color:turquoise3'>kid</span></strong></b> versus <strong><span style='color:salmon2'>not kid</span></strong></b> recommended trails")+
  theme_minimal()+
  theme(plot.title.position = "plot", 
        panel.grid.major.x = element_blank(), 
        panel.grid.minor.x = element_blank(),
        plot.title = element_markdown(family = "mono", face = "bold", size = 10))
```

After getting the sense that kid friendly trails tend to be shorter, we decided to look more into each park and see if there might be one that is more attractive to families than another. To do this, we looked at the number of kid friendly trails and the average rating those trails got. The results are summarized below. 

```{r, echo = FALSE}
kids_proportion %>%
  mutate(area_name = word(area_name, 1, -3)) %>%
  ggplot(aes(y=fct_reorder(area_name, kids_yes), x=kids_yes))+
  geom_col(aes(fill = kids_avg_rating)) + 
  labs(y="", x="", 
       title = "Number of trails that are <strong><span style='color:dodgerblue3'>kid friendly</span></strong></b>",
       fill = "Avg <strong><span style='color:dodgerblue3'>kid friendly</span></strong></b> \ntrail rating")+
  scale_x_continuous(expand = c(0,0))+
  theme_minimal()+
  theme(plot.title = element_markdown(family = "mono", face = "bold"),
        legend.title = element_markdown(family = "mono", size = 7),
        plot.title.position = "plot", 
        panel.grid.major.x = element_blank(), 
        panel.grid.minor.x = element_blank(), 
        panel.grid.major.y = element_blank(), 
        panel.grid.minor.y = element_blank())
```

From this graph we can see both Yellowstone and Acadia have close to 80 kid recommended trails...it might take quite a few family vacations to cover all of those :) On the lower end are Grand Teton and Grand Canyon National with 24 and 27 kid recommended trails each. The graph is colored by average kid friendly trail rating. While Grand Teton has the fewest number of kid friendly trails - though still plenty for a family vacation with over 52 miles of these trails - the average trail rating is the highest. Joshua Tree has the lowest average kid recommended trail rating at around 3.8/5. 

To wrap up our discussion of the national parks and kid friendly trails, we recommended two family friendly trails from each of the ten parks we looked at. To choose these parks, we looked for trails within our data that have a high number of reviews and at least 4.5 stars. The table below highlights some details of these trails including a few of their unique and/or exciting features. 

```{r, echo = FALSE}
recommend_trails <- national_parks %>%
  filter(name %in% c("Peregrine Peak via Alum Cave Bluffs Trail", 
                     "Clingmans Dome Observation Tower Trail", 
                     "Grand Prismatic Hot Spring", 
                     "Mystic Falls, Fairy Creek and Little Firehole Loop", 
                     "Zion Canyon Overlook Trail", 
                     "The Watchman Trail", 
                     "Emerald Lake Trail", 
                     "Dream Lake Trail", 
                     "Hidden Falls Trail", 
                     "String Lake Trail", 
                     "Shoshone Point Trail", 
                     "Grand Canyon Rim Trail", 
                     "Ledges Trail", 
                     "Brandywine Gorge Trail", 
                     "Cadillac North Ridge Trail", 
                     "Jordan Pond Full Loop Trail", 
                     "Hoh Rain Forest Hall of Moss", 
                     "Marymere Falls Trail", 
                      "Ryan Mountain Trail", 
                    "Hidden Valley Nature Trail")) %>%
  select(area_name, name, length_miles, avg_rating, num_reviews)%>%
  head(20) %>%
  mutate(`noteable features` = c("views, wildlife", "views, wildlife", "hot springs, river", "hot springs, river, waterfall", "cave, trail running", "river, camping", "river, fishing, snowshoeing", "skiing, snowshoeing, rock climbing", "lake, beach, rock climbing, camping", "ada, beach, strollers, birding", "ada, strollers, birding, trail running", "ada, strollers", "historic site, trail running", "historic site, waterfall, river, fishing", "trail running, birding, wildlife", "lake, birding, trail running", "ada, strollers, camping, birding", "lake, waterfall, birding", "wildlife, views", "views, rock climbing, trail running")) %>%
  rename(Park = area_name, Trail = name, `Length (miles)` = length_miles, `Avg rating` = avg_rating, `Number reivews` = num_reviews, `Noteable features` = `noteable features`) %>%
  arrange(Park, Trail)

trail_recs_table <- recommend_trails %>%
  gt() %>%
  tab_header(
    title = md("**Exciting trails for kids**"),
    subtitle = md("**Two kid recommended trails from each of the top 10 most visited national parks**")
  ) %>%
  tab_style(
    style = cell_fill(color = "lightgoldenrodyellow"),
    locations = cells_body(
      rows = Park %in% c("Acadia National Park", "Grand Canyon National Park", "Great Smoky Mountains National Park", "Olympic National Park", "Yellowstone National Park"))) %>%
  tab_style(
    style = cell_fill(color = "lightcyan2"),
    locations = cells_body(
      rows = Park %!in% c("Acadia National Park", "Grand Canyon National Park", "Great Smoky Mountains National Park", "Olympic National Park", "Yellowstone National Park"))) %>%
  tab_options(
    table.font.size = px(13L)
  )
trail_recs_table
```

We hope you enjoy your next trip to one of the many beautiful US national parks! 

